### 2020-12-20 paper/abstract.tex
Enforced 88 character limit.
small adjustments.

\noindent Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs. Epidemiological models have a hard time predicting the effects of such fine-grained policies. We propose a novel simulation-based model to address this shortcoming. We build on state-of-the-art agent-based simulation models but replace the way contacts between susceptible and infected people take place. Firstly, we allow for heterogeneity in the types of contacts (e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we strictly separate the number of contacts from the probabilities that a contact leads to an infection. The number of contacts changes with social distancing policies, the infection probabilities remain invariant. This allows us to model many types of fine-grained policies that cannot easily be incorporated into other models. To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.

>>>>

\noindent Governments worldwide are adopting nuanced policy measures to reduce the
number of Covid-19 cases and to balance social and economic costs. Epidemiological
models have a hard time predicting the effects of fine-grained policies. We propose a
novel simulation-based model to address this shortcoming. We build on state-of-the-art
agent-based simulation models but replace the way contacts between susceptible and
infected people take place. Firstly, we allow for heterogeneity in the types of contacts
(e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we
strictly separate the number of contacts from the probabilities that a contact leads to
an infection. The number of contacts changes with social distancing policies, the
infection probabilities remain invariant. This allows us to model many types of targeted
policies that cannot be incorporated into other models. To validate our model, we show
that it can predict the effect of the German November lockdown even if no similar policy
has been observed during the time period that was used to estimate the model parameters.

### 2020-12-20 paper/introduction.tex
small adjustments.

The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
across the world. As a result, many countries were able to reduce the number of positive
cases and could relax restrictions imposed on their societies. At the same time, the
social and economic costs were enormous. In the second wave, countries are trying to
implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in
November. Some businesses are closed and certain leisure activities are prohibited but
schools and nurseries stay open.

Current epidemiological models have not been designed to predict the effect of such
fine-grained policies. These models need to be extended for each new policy proposal and
cannot be easily adapted to fast changing environments. This report describes a model
that has been designed from the ground up to predict the effects of contact reducing
policies in real time. At the time of this writing, it has the following features:

\begin{enumerate}
    \item At the core of the model, people meet people based on a matching algorithm. We
          distinguish various types of contacts. Currently, these are households,
          leisure activities, schools, nurseries and several types of contacts at the
          workplace. Contact types can be random or recurrent and vary in frequency.

@@

    \item Infection probabilities vary across contact types, but are invariant to
          policies which reduce contacts.

    \item The model achieves a good fit on German infection and fatality rate data even
          if only the infection probabilities are fit to the data and the remaining
          parameters are calibrated from the medical literature and datasets on contact
          frequencies.

    \item High quality Python code for the model is freely available on Github, well
          documented and very flexible\footnotemark. We are actively looking for
          researchers who want to use our model for their projects.
\end{enumerate}

@@

@@

More details about the German model as well as applications to currently discussed
German policies can be found in a companion paper
(\citet{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new
report.}

>>>>

The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions
across the world. Many countries were able to contain the spread of the virus and to
relax restrictions during midyear. At the same time, the social and economic costs were
enormous. In the second wave, countries are trying to implement more nuanced policies.
For example, Germany imposed a ``lockdown-light'' in November. Some businesses are
closed and certain leisure activities are prohibited but educational facilities remain
open.

Current epidemiological models have not been designed to predict the effect of
fine-grained policies. The models need to be extended for each new policy proposal and
cannot be easily adapted to fast changing environments. This report describes a model
that has been designed from the ground up to predict the effects of contact reducing
policies in real time. It has the following features:

\begin{enumerate}
    \item At the core of the model, people meet people based on a matching algorithm. We
          distinguish various types of contacts. The contact types are households,
          leisure activities, schools, preschools, and nurseries and several types of
          contacts at the workplace. Contact types can be random or recurrent and vary
          in frequency.

@@

    \item Infection probabilities vary across contact types and reflect properties of
          the contact like the location (indoor/outdoor) and the kind of interaction
          (duration, physical contact). The probabilities are independent from the
          number of contacts and, thus, policy-invariant.

    \item The model achieves a good fit on German data of infection rates even if only
          the infection probabilities are fit to the data and the remaining parameters
          are calibrated from the medical literature and datasets on contact
          frequencies.

    \item High quality Python code for the model is freely available on Github, well
          documented and very flexible\footnotemark. We are actively looking for
          researchers who want to use our model for their projects and employ it in
          different contexts.
\end{enumerate}

@@

More details about the German model as well as applications to currently discussed
German policies can be found in \citet{Dorn2020a}.

@@

### 2020-12-20 paper/model/matching.tex
small adjustments.

The matching process for recurrent contacts is then extremely simple: On each simulated
day, every person who does not stay home meets all other group members who do not stay
home. The assumption that all group members have contacts with all other group members
is not fully realistic, but seems like a good approximation to reality, especially in
light of the suspected role of aerosol transmission for Covid-19 (\citet{Morawska2020,
Anderson2020}).

>>>>

The matching process for recurrent contacts is then extremely simple: On each simulated
day, every person who does not stay home meets all other group members who do not stay
home. The assumption that all group members have contacts with all other group members
is not fully realistic, but seems like a good approximation to reality, especially in
light of the suspected role of aerosol transmission for Covid-19 \citep{Morawska2020,
Anderson2020}.
